---
title: "Web-Scraping NBA"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import the necessary packages
```{r message=FALSE,warning=FALSE}
library(tidyverse) # manipulation, visualisation, tidying data
library(janitor) # cleaning the data
library(rvest) # scrapping the web data
#install.packages("hablar") # converts data type
library(hablar)
library(flextable)
```

## Declaring the url of the table to be imported
```{r message=FALSE,warning=FALSE}
url = "https://www.basketball-reference.com/allstar/NBA-allstar-career-stats.html"
# create a var pageobj to read the url address and transform the content into a data frame
pageobj = read_html(url, as.tibble = T)

# create a variable ALLStarPlayes that is holding the content of the first table from the webpage we imported anterior.
AllStarsPlayers = pageobj %>% 
  html_nodes("table") %>% .[[1]] %>% 
  html_table(fill = T)

view(AllStarsPlayers)
glimpse(AllStarsPlayers)
```


```{r message=FALSE,warning=FALSE}
# Notice for All-Star Player stats there is a single data table on the website we specified. Therefore, as shown above we referenced the page object as [[1]] indicating it is the first (in this case only) data table on the site.
```

## Data Cleaning: Updating Variable Names
```{r message=FALSE,warning=FALSE}
# The janitor package offers two functions that help us with the cleaning proocess:
# Converts all row names to follow tidyverse style requirements
# column names are changed from capital letters to small letters
AllStarsPlayers = AllStarsPlayers %>% 
  row_to_names(row_number = 1, remove_row = T, remove_rows_above = F) %>% 
  clean_names()

view(AllStarsPlayers)
names(AllStarsPlayers)
```

```{r message=FALSE,warning=FALSE}
# In AllStarPlayers ast: is the total number of assists across all games
# In AllStarPlayers ast2: is the average number of assists per game
# In ONeal ast: is the average number of assists per game
# To avoid confussion we have to create column names that are more descriptive
AllStarsPlayers = AllStarsPlayers %>% 
  rename("mins_total" = "mp",
         "pts_total" = "pts",
         "trb_total" = "trb",
         "ast_total" = "ast",
         "mins_per_game" = "mp_2",
         "pts_per_game" = "pts_2",
         "trb_per_game" = "trb_2",
         "ast_per_game" = "ast_2")

names(AllStarsPlayers)
```
## Data Cleaning: Remove Subheadings
```{r message=FALSE,warning=FALSE}
# The AllStarPlayers dataset has multiple rows that repeat the variable heading names. This is a useful feature when scrolling through large datasets online, as there is always a row displayed on the screen to reference.

get_dupes(AllStarsPlayers) %>% flextable()
```


```{r message=FALSE,warning=FALSE}
AllStarsPlayers = AllStarsPlayers %>% 
  filter(
    player != "Player" & player != ""
  )

view(AllStarsPlayers)
```

## Create New Variables
```{r message=FALSE,warning=FALSE}
# verify the the feature characteristics
glimpse(AllStarsPlayers)
# As one can see all the columns has been imported by rvest as characters we have now to work on converting them into integers and numeric
AllStarsPlayers = AllStarsPlayers %>% 
  convert(int("g", "gs", "mins_total", "fg", "fga", "x3p", "x3pa", "x2p",    
      "x2pa", "ft", "fta", "orb", "drb", "trb_total", "ast_total",  
      "stl", "blk", "tov", "pf", "mins_per_game", "pts_per_game",    
      "trb_per_game", "ast_per_game"),
      num("fg_percent", "x3p_percent", "x2p_percent", "ft_percent"))
# evaluate the transformation
glimpse(AllStarsPlayers)
```


```{r message=FALSE,warning=FALSE}
# Other data sets, especially those for individual players, calculate the stats as averages per-game. Having variables on the same scale allows us to merge and compare data.
# Converting the totals to per-game averages can be done using traditional math operators.

AllStarsPlayers = AllStarsPlayers %>% 
  mutate(
    fg_per_game = fg / g,
    fga_per_game = fga / g,
    x3p_per_game = x3p / g
  ) %>% 
  rename(
    "fg_total" = "fg",
    "fga_total" = "fga",
    "x3p_total" = "x3p"
  )

names(AllStarsPlayers)
```
## Import the O'Neal Data set
```{r message=FALSE, warning=FALSE}
# put the link into a variable
url1 = "https://www.basketball-reference.com/players/o/onealsh01.html"
# read the url link with rvest
pageobj1 = read_html(url1, as.tibble = T)
# extract the necessary table by specifying the nr of the table on the web page
Oneal = pageobj1 %>% html_nodes("table") %>% .[[1]] %>% html_table()
# get a glimpse of the dta imported
glimpse(Oneal)
view(Oneal)
```

## Select only the evident rows
```{r message=FALSE,warning=FALSE}
# use slice function to sebset a data set
Oneal = Oneal %>% 
  slice(1:21) %>% 
  clean_names()
# check integrity of the data frame 
view(Oneal)
glimpse(Oneal)
```
## Add the column Player to the Oneal data frame
```{r message=FALSE,warning=FALSE}
Oneal = Oneal %>% 
  mutate(
    player = "Shaquille O'Neal"
  )

AllStarsPlayers = AllStarsPlayers %>% 
  mutate(
    season = "AllStar"
  )
```


## Compare the names of the two data frames
```{r}
# Instead of individually comparing the files we can use the janitor function compare_df_cols_same().
compare_df_cols_same(AllStarsPlayers, Oneal, bind_method = "rbind")

```
## Remove the columns from AllStarPlayers that are not present in Oneal 
```{r}
AllStarsPlayers = AllStarsPlayers %>% 
  select(-ast_total, -ast_per_game, -fg_total, -fga_total, -pf_total, -stl_total,  
         -tov_total, -x2p_total, -x2pa_total, -x3pa_total, 
         -x3p_total, -mins_total, -pts_total, -trb_total)
```











































































